---
title: "Homework 1"
author: "Jake Fenley"
date: "9/16/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# How to Submit
1) output as github_document and push to github
Then output as html_document and upload html_doc to canvas
2) Include link to github in your html document

## [Link to Github](https://github.com/jpFenley/stat433-hw1)

### Import necessary libraries
```{r}
library(rvest)
library(stringr)
library(plyr)
library(kableExtra)

```

### Read in HTML
We can read in the HTMl from the website and create a data frame to hold the information.
```{r}
html <- read_html("https://guide.wisc.edu/faculty/")
people <- html_nodes(html, '.uw-people')
faculty = data.frame(Name=character(),
                     Title = character(),
                     Area=character(),
                     School = character(),
                     stringsAsFactors = FALSE)

```

### Parse Data
We begin by iterating through all the letters of the alphabet. We remove unnecessary HTML at the beginning and end of each table. Then we can split 
the information by each person. For each person, we split the data by break symbols to obtain all of the fields. We update the table each time.
```{r}

for (letter in people) {
  letter <- as.character(letter)
  letter <- str_replace(letter, "<ul class=\"uw-people\">\n<li><p>", "")
  letter <- str_replace(letter, "<br></p></li>\n</ul>\n", "")
  letter <- str_replace(letter, "<br></p></li>\n</ul>", "")
  names <- unlist(strsplit(letter, "<br></p></li>\\n<li><p>"))
  for (p in names){
    faculty <- rbind(faculty, unlist(strsplit(p, "<br>")))
  }
}
names(faculty) <- c("Name", "Title", "Area", "Degree")

```

### Checking Data
We will print the first couple rows of the table.

```{r}
head(faculty)

```

### Creating a Plot
We wish to plot the count of how many last names start with each letter. We assign a new column to the data frame with this letter. We then count how many each one appears, and plot is as a bar plot. 
```{r}
faculty$Starts <- str_sub(faculty$Name, 1, 1)

```

```{r, fig.width=9}
cnts <- table(faculty$Starts)
barplot(cnts, main="Most Common Last Name Starting Letter",
        xlab = "Starting Letter",
        ylab = "Frequency")

```